DATA 621---Business Analytics and Data Mining

### 2020-09-10 HW1/HW1_AF.Rmd
more updates

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}

@@

```{r, include = FALSE}

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}

@@

ggplot(data = negSet %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
  facet_wrap(~ STAT, scales = "free")

@@

Many of these stats *do* seem to have an effect on the number of wins. Most notably,
as the number of base hits and walks increase, so do the number of wins. We see
weaker positive relationships for homeruns, doubles, triples, and stolen bases.
This makes sense when we think about it -- these things tend to happen less often in
games than pure base hits and walks, so they don't have as much of an effect on winning. 
Double plays and batters hit don't appear to have an effect on the number of wins either. 
Once again, this intuitively makes sense because they are less likely to happen in a game.

One thing to note is the number of strikeouts compared to the number of wins. We can see
that there are a few outliers -- abnormally high numbers of strikeouts in a season. This
should be taken with caution, as they don't represent a typical season's stats.

```{r, echo = FALSE, message = FALSE, warning = FALSE}

@@

Some of the provided statistics are well-balanced in the sense that there are
very few (or no) extreme values. **Homeruns by batters**, **Strikeouts by** 
**batters**, and **Homeruns allowed** are examples of this. Aside from

```{r, echo = FALSE, message = FALSE, warning = FALSE}

@@

# sapply(train_set %>% select(-INDEX, -TARGET_WINS), function(x){length(boxplot.stats(x)$out)})

>>>>

```{r wins, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}

@@

```{r pctNull, include = FALSE}

@@

```{r negFactors, echo = FALSE, message = FALSE, warning = FALSE}

@@

ggplot(data = negSet %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
  facet_wrap(~ STAT, scales = "free")
```

@@

Many of these stats *do* seem to have an effect on the number of wins, most notably,
base hits and walks. We see weaker positive relationships for homeruns, doubles, triples, 
and stolen bases. This makes sense when we think about it; these things tend to happen 
less often in games than pure base hits and walks, so they don't have as much of an 
effect on winning. Finally, double plays and batters hit don't appear to have any 
correlation with the number of wins. Once again, this intuitively makes sense because 
they are less likely to happen in a game.

One thing to note is the number of strikeouts compared to the number of wins. We can see
that there are a few outliers (abnormally high numbers of strikeouts in a season). This
should be taken with caution, as they don't represent a typical season's stats.

```{r posFactors, echo = FALSE, message = FALSE, warning = FALSE}

@@

Some of the provided statistics are well-balanced in the sense that there are
very few (or no) extreme values. **Homeruns by batters**, **Strikeouts by** 
**batters**, and **Homeruns allowed** are examples of this.

```{r summaryInfo, include = FALSE}

summary(train_set)
```

```{r fewOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}

@@

```
Some things to note about each of these statistics:

* Homeruns allowed (average = ~100/year) and Homeruns by batters (average = ~106/year) 
have a very similar mid-range distribution (50% of the data lies between ~50 and 150). 
The slight difference in average stats means that teams tend to have a higher number
of homeruns than the opposition team. 
* The only thing that stands out about strikeouts by batters (average = ~736/year) is 
how nearly perfectly normal it is. 50% of the data is between about 500 and 1000 and 
there are absolutely no outliers in the dataset! This means that there were no surprisingly
high or low seasons.

Conversely, some of the stats have a very high number of outliers, indicating that there
are some seasons with some abnormally high or low values. **Walks by batters**, **Stolen**
**bases**, **Hits allowed**, and **Errors** are examples of this.

```{r manyOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5, fig.align = 'center'}

set2 <-  train_set %>%
  select('Walks by batters' = TEAM_BATTING_BB, 
         'Stolen bases' = TEAM_BASERUN_SB,  
         'Hits allowed' = TEAM_PITCHING_H,
         'Errors' = TEAM_FIELDING_E) %>%
  gather("stat", "value") %>%
  filter(complete.cases(.) == TRUE)

vals2 <- ggplot(set2 , aes(x=stat, y=value)) + 
    geom_boxplot() +
  labs(title="",x="", y = "Value") +
  facet_wrap(~ stat, scale = 'free')

vals2

#sapply(train_set %>% select(-INDEX, -TARGET_WINS), function(x){length(boxplot.stats(x)$out)})

```
Some things to note about each of these statistics:
